An assessment of photometric redshift PDF performance in the context of LSST

### 2018-09-04 paper/tex/results.tex
add scotts rule smoothing to Figure 4
add/modify text describing Fig 4 update

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{fig/PITANDQQplot_12codes_crop.jpg}
\caption{Summary plots for all eleven photo-$z$ codes illustrating performance for the interim posterior statistics. The top panel of each pair shows both the Quantile-Quantile (QQ) plot (red) and the histogram of PIT values (blue).  The desired behavior is a QQ plot that matches the diagonal dashed line, and a PIT histogram that matches a uniform distribution matching the thin horizontal black line.  The bottom panel of each pair shows the difference between the QQ quantile and the diagonal, illustrating departure from the desired performance.  Histograms with an overabundance of PIT values at the centre of the distribution indicate $p(z)$ distributions that are overly broad, while an excess of values at the extrema indicate $p(z)$ distributions that are overly narrow.  Values of PIT=0 and PIT=1 indicate ``catastrophic failures'' where the true redshift is completely outside the support of $p(z)$.  Asymmetric features are indicative of systematic bias in the redshift predictions.  A variety of behaviors are evident, and specific details are discussed in the text.}
\label{fig:pitqq}
\end{figure*}

Fig.~\ref{fig:pz_examples} Shows the $p(z)$ produced by each of our eleven photo-$z$ codes for four example galaxies which exemplify some prominent cases that arise when estimating photo-$z$ PDFs: a narrow, unimodal redshift solution, a broader unimodal solution, a bimodal distribution, and a complex, multimodal distribution.
The red vertical line represents the true redshift of the individual galaxy, and the blue curve represents the redshift probability.
Several features are obvious even in these illustrative examples.
\textsc{ANNz2, METAPhoR, NN,} and \textsc{SkyNet} all show an excess of small-scale features, which appear to be print-through of the underlying training set galaxies.
\textsc{GPZ} (in its current implementation), on the other hand, always produces a single Gaussian, which broadens to cover the multi-modal redshift solutions seen in other codes.

As stated in Section~\ref{sec:metrics}, $p(z)$ is parameterized as $\approx 200$ piecewise constant bins covering $0<z<2$ for all eleven codes, giving a grid size of roughly $\delta z = 0.01$ for each code.
A piecewise constant grid was a natural choice for some photo-$z$ codes, for instance most template-based codes compute likelihoods on a fixed grid.
In contrast, FlexZBoost, for example, can return estimates on any grid without compression errors as it’s a basis expansion method where only the expansion coefficients need to be stored.
Codes with a native output format other than the shared piecewise constant binning scheme (or one that can be losslessly converted to it) may suffer from loss of information when converting to it, which could artificially favor some codes over others.

@@

Examination of the PIT histograms and QQ plots shows that there are fairly generic issues with the width of $p(z)$ uncertainties: \textsc{Delight, NN, SkyNet} and \textsc{TPZ} all show a PIT histogram with an dearth of low values and and an excess of high values, signs that, on average, their $p(z)$ are more broad than the true distribution of redshifts.  \textsc{METAPhoR} shows the opposite trend, indicating the the $p(z)$  are more narrow than the distributions given by the true redshifts.  In all of these code cases there is a free parameter or bandwidth that can be used to tune uncertainties.  The sensitivity of multiple codes to this bandwidth choice emphasizes the fact that great care must be taken in setting user-defined parameters in photo-$z$ codes, even in the presence of representative training/validation data.  for \textsc{FlexZBoost} the ``sharpening'' parameter (described in Section~\ref{sec:flexzboost}) plays a key role in improving the results, resulting in a QQ plot that is very nearly diagonal.  A similar sharpening procedure could be beneficial for several codes.
Interestingly, the three purely template-based codes, \textsc{BPZ, EAZY}, and \textsc{LePhare}, show relatively well behaved $p(z)$ statistics (albeit with some bias), which may indicate that the likelihood estimation with representative templates is accurately capturing the uncertainties on individual redshifts.

@@

Fig.~\ref{fig:nz} shows the stacked $\hat{N}(z)$ distribution compared to the true redshift distribution $N'(z)$ for all tested codes.  The red line indicates the summed $p(z)$ for each code, while the blue line shows the true redshift distribution smoothed via kernel density estimation (KDE), with a bandwidth chosen via Scott's rule \citep{Scott:1992}.  While Scott's rule is used to display $N'(z)$ in the figure, all quantitative statistics are computed via the empirical CDF, and are thus unaffected by bandwidth/smoothing choice.
Several of the codes show an excess at $z \sim 1.4$, particularly the template-based codes \textsc{BPZ}, \textsc{EAZY}, and \textsc{LePhare}.
This is likely due to the $4000$ angstrom break passing through the gap between the $z$ and $y$ filters.
This feature is one of the most prominent in individual galaxy p(z), and is readily seen in the point-estimate plots shown in Fig.~\ref{fig:pz_pointestimates} and described in the Appendix.
Several of the machine learning based codes appear to be over-trained, adding excess galaxy probability to the redshift peaks and missing probability in the troughs.
Given that our training data is drawn from the same galaxy population as the test set, and our data has prominent peaks in $N'(z)$, perhaps it is not unexpected that such overtraining occurs.
A more extensive training/validation set might allow for a better choice of smoothing parameters in individual codes that would avoid such overtraining.

%\aim{Move the text on Scott's Rule to Discussion?}
%Scott's rule \citep{Scott:1992} was used to determine the smoothing bandwidth of the true redshift sample: that is, a single smoothing scale was chosen to represent the true redshift sample for all eleven codes.

As with the $p(z)$ values in Figure~\ref{fig:pitqq}, different levels of substructure are obvious for the different codes.
While Scott's rule provides a relatively good general smoothing scale to represent the true $N'(z)$, there are smaller scale fluctuations: while \textsc{FlexZBoost} and \textsc{CMNN} appear somewhat discrepant in Fig.~\ref{fig:nz}, they are actually the two most accurate in terms of their quantitative measurements.
%\textsc{SkyNet} also shows excess small-scale structure, but also shows significant overestimate of the high redshift galaxy population.
Interestingly, while \textsc{ANNz2} shows an abundance of small scale structure in individual $p(z)$ measurements (see Fig.~\ref{fig:pz_examples}), the summed $\hat{N}(z)$ is rather smooth, where the small scale features average out.  This is not the case for the two other codes that show an abundance of substructure in their individual $p(z)$: both \textsc{CMNN} and \textsc{SkyNet} show small scale features both in $p(z)$ and $\hat{N}(z)$.  For \textsc{CMNN} the $p(z)$ are a simply a weighted histogram of all spectroscopic training galaxies in nearby colour space with no smoothing applied, so the substructure is not unexpected.  The PIT histogram and shape of the QQ plot in Figure~\ref{fig:pitqq} show that \textsc{CMNN} is producing $p(z)$ that are overly broad, additional smoothing of the $p(z)$ would exacerbate this problem.  While the $\hat{N}(z)$ plot shows more small scale features than other codes, these features are actually representative of real structure in the true $N'(z)$, as evidenced by the very good metric scores for \textsc{CMNN}.  \textsc{SkyNet} $p(z)$ were also not smoothed: while previous implementations of the code such as \citet{Sanchez:14} and \citet{Bonnett:15} (see Appendix C.3) implement a ``sliding bin'' smoothing, no such procedure was used in this study.  In addition to excess substructure, \textsc{SkyNet} shows an obvious redshift bias, evident both visually in Figure~\ref{fig:nz} and in the first moment of N(z) listed in Table~\ref{tab:moments}, where it is clearly an outlier.  \textsc{SkyNet} employed a method where a random sample of training galaxies was chosen, but there was no test that the subset was completely representative of the overall redshift distribution.  Also unlike \citet{Bonnett:15}, no effort was made to add extra weight to more rare low and high redshift galaxies.  Either of these decisions could be the cause of the bias seen in our results.  Future runs of \textsc{SkyNet} will explore these implementation choices and their effects.

@@

\begin{table}
\setlength{\tabcolsep}{2pt}
\begin{center}
\caption{Root-Mean-Square-Error (RMSE) statistics for the eleven photo-z codes for both PIT and $\hat{N}(z)$ distributions.}\label{tab:rmse}
\begin{tabular}{lcc}
\hline
\hline
Root-Mean-Square-Error \\
(RMSE) statistics \\
\hline
Photo-z Code & PIT RMSE & N(z) RMSE\\
\hline
\textsc{ANNz2}      & 0.019 & 0.0054\\
\textsc{BPZ}        & 0.032 & 0.0050\\
\textsc{Delight}    & 0.111 & 0.0056\\
\textsc{EAZY}       & 0.054 & 0.0102\\
\textsc{FlexZBoost} & 0.021 & 0.0022\\
\textsc{GPz}        & 0.027 & 0.0042\\
\textsc{LePhare}    & 0.028 & 0.0062\\
\textsc{METAPhoR}   & 0.064 & 0.0081\\
\textsc{CMNN}         & 0.108 & 0.0009\\
\textsc{Skynet}     & 0.054 & 0.0144\\
\textsc{TPZ}        & 0.082 & 0.0031\\
\hline
\trainz		& 0.0025 & 0.0013\\
\end{tabular}
\end{center}
\end{table}

>>>>

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{fig/PITANDQQplot_12codes_crop.jpg}
\caption{Summary plots for all twelve photo-$z$ codes illustrating performance for the interim posterior statistics. The top panel of each pair shows both the Quantile-Quantile (QQ) plot (red) and the histogram of PIT values (blue).  The desired behavior is a QQ plot that matches the diagonal dashed line, and a PIT histogram that matches a uniform distribution matching the thin horizontal black line.  The bottom panel of each pair shows the difference between the QQ quantile and the diagonal, illustrating departure from the desired performance.  Histograms with an overabundance of PIT values at the centre of the distribution indicate $p(z)$ distributions that are overly broad, while an excess of values at the extrema indicate $p(z)$ distributions that are overly narrow.  Values of PIT=0 and PIT=1 indicate ``catastrophic failures'' where the true redshift is completely outside the support of $p(z)$.  Asymmetric features are indicative of systematic bias in the redshift predictions.  A variety of behaviors are evident, and specific details are discussed in the text.}
\label{fig:pitqq}
\end{figure*}

Fig.~\ref{fig:pz_examples} Shows the $p(z)$ produced by each of our twelve photo-$z$ codes for four example galaxies which exemplify some prominent cases that arise when estimating photo-$z$ PDFs: a narrow, unimodal redshift solution, a broader unimodal solution, a bimodal distribution, and a complex, multimodal distribution.
The red vertical line represents the true redshift of the individual galaxy, and the blue curve represents the redshift probability.
Several features are obvious even in these illustrative examples.
\textsc{ANNz2, METAPhoR, NN,} and \textsc{SkyNet} all show an excess of small-scale features, which appear to be print-through of the underlying training set galaxies.  For example, in \textsc{CMNN} the $p(z)$ are a simply a weighted histogram of all spectroscopic training galaxies in nearby colour space with no smoothing applied, so the substructure is due to the finite number of neighbours, and is not unexpected.
\textsc{GPZ} (in its current implementation), on the other hand, always produces a single Gaussian, which broadens to cover the multi-modal redshift solutions seen in other codes.
%Interestingly, while \textsc{ANNz2} shows an abundance of small scale structure in individual $p(z)$ measurements (see Fig.~\ref{fig:pz_examples}), the summed $\hat{N}(z)$ is rather smooth, where the small scale features average out.  This is not the case for the two other codes that show an abundance of substructure in their individual $p(z)$: both \textsc{CMNN} and \textsc{SkyNet} show small scale features both in $p(z)$ and $\hat{N}(z)$.

As stated in Section~\ref{sec:metrics}, $p(z)$ is parameterized as $\approx 200$ piecewise constant bins covering $0<z<2$ for all twelve codes, giving a grid size of roughly $\delta z = 0.01$ for each code.
A piecewise constant grid was a natural choice for some photo-$z$ codes, for instance most template-based codes compute likelihoods on a fixed grid.
In contrast, FlexZBoost, for example, can return estimates on any grid without compression errors as it’s a basis expansion method where only the expansion coefficients need to be stored.
Codes with a native output format other than the shared piecewise constant binning scheme (or one that can be losslessly converted to it) may suffer from loss of information when converting to it, which could artificially favor some codes over others.

@@

Examination of the PIT histograms and QQ plots shows that there are fairly generic issues with the width of $p(z)$ uncertainties: \textsc{Delight, CMNN, SkyNet} and \textsc{TPZ} all show a PIT histogram with an dearth of low values and and an excess of high values, signs that, on average, their $p(z)$ are more broad than the true distribution of redshifts.  \textsc{METAPhoR} shows the opposite trend, indicating the the $p(z)$  are more narrow than the distributions given by the true redshifts.  In all of these code cases there is a free parameter or bandwidth that can be used to tune uncertainties.  The sensitivity of multiple codes to this bandwidth choice emphasizes the fact that great care must be taken in setting user-defined parameters in photo-$z$ codes, even in the presence of representative training/validation data.  for \textsc{FlexZBoost} the ``sharpening'' parameter (described in Section~\ref{sec:flexzboost}) plays a key role in improving the results, resulting in a QQ plot that is very nearly diagonal.  A similar sharpening procedure could be beneficial for several codes.
Interestingly, the three purely template-based codes, \textsc{BPZ, EAZY}, and \textsc{LePhare}, show relatively well behaved $p(z)$ statistics (albeit with some bias), which may indicate that the likelihood estimation with representative templates is accurately capturing the uncertainties on individual redshifts.

@@

Fig.~\ref{fig:nz} shows the stacked $\hat{N}(z)$ distribution compared to the true redshift distribution $N'(z)$ for all tested codes.  The red line indicates the summed $p(z)$ for each code, while the blue line shows the true redshift distribution. All distributions are smoothed via kernel density estimation (KDE) with a common bandwidth chosen via Scott's rule \citep{Scott:1992} in order to minimize differences in small-scale features and make for a more uniform comparison between codes.  While Scott's rule is used to display $N'(z)$ in the figure, all quantitative statistics are computed via the empirical CDF, and are thus unaffected by bandwidth/smoothing choice.
%New text for the new figure
As expected, \textsc{TrainZ} is in excellent agreement with true redshift distribution: as the training sample is selected from the same underlying distribution as the test set, the redshift distributions are identical, up to Poisson fluctuations due to the finite number of sample galaxies.  \textsc{CMNN} is also in excellent agreement for similar reasons: with a representative training sample of galaxies spanning the colour-space, the sum of the colour-matched neighbour redshifts should return the true redshift distribution. \textsc{FlexZBoost} and \textsc{TPZ} also show very good agreement, with only slight departures, with an over/under-prediction in the high redshift tail of $\hat{N}(z)$ evident around $z\sim1.4$.  In fact, several of the other codes show an excess at $z \sim 1.4$, particularly the template-based codes \textsc{BPZ}, \textsc{EAZY}, and \textsc{LePhare}.
This is likely due to the $4000$ angstrom break passing through the gap between the $z$ and $y$ filters, resulting in a drastic change in $z-y$ colour for galaxies in this redshift range.  The relative dearth of strong features blue-ward of the 4000 angstrom break seems to cause problems for the redshift range that corresponds to the break passing between the two reddest filters bandpasses of a survey.
The $z\sim1.4$ feature is one of the most prominent sources of larger uncertainty in individual galaxy $p(z)$, and is readily seen in the point-estimate plots shown in Fig.~\ref{fig:pz_pointestimates} and described in the Appendix.
%end of new text (and slightly modified text)
Two of the machine learning based codes, \textsc{ANNz2} and \textsc{METAPhoR}, appear to be over-trained, adding excess galaxy probability to the redshift peaks and missing probability in the troughs.
Given that our training data is drawn from the same galaxy population as the test set, and our data has prominent peaks in $N'(z)$, perhaps it is not unexpected that such overtraining occurs in some codes, though the fact that it does not occur in all training-based codes indicates that it may be due to specifics of the implementations of \textsc{ANNz2} and \textsc{METAPhoR}.
A more extensive training/validation set might allow for a better choice of smoothing parameters in individual codes that would avoid such overtraining for these particular codes.
%\aim{Move the text on Scott's Rule to Discussion?}
%Scott's rule \citep{Scott:1992} was used to determine the smoothing bandwidth of the true redshift sample: that is, a single smoothing scale was chosen to represent the true redshift sample for all twelve codes.
%As with the $p(z)$ values in Figure~\ref{fig:pitqq}, different levels of substructure are obvious for the different codes.
%While Scott's rule provides a relatively good general smoothing scale to represent the true $N'(z)$, there are smaller scale fluctuations: while \textsc{FlexZBoost} and \textsc{CMNN} appear somewhat discrepant in Fig.~\ref{fig:nz}, they are actually the two most accurate in terms of their quantitative measurements.
%\textsc{SkyNet} also shows excess small-scale structure, but also shows significant overestimate of the high redshift galaxy population.
\textsc{SkyNet} shows an obvious redshift bias, evident both visually in Figure~\ref{fig:nz} and in the first moment of N(z) listed in Table~\ref{tab:moments}, where it is clearly an outlier.  \textsc{SkyNet} employed a method where a random sample of training galaxies was chosen, but there was no test that the subset was completely representative of the overall redshift distribution.  Unlike the previous implementation of \textsc{SkyNet} in \citet{Bonnett:15}, no effort was made to add extra weight to more rare low and high redshift galaxies.  Either of these decisions could be the cause of the bias seen in our results.  Future runs of \textsc{SkyNet} will explore these implementation choices and their effects.

@@

\begin{table}
\setlength{\tabcolsep}{2pt}
\begin{center}
\caption{Root-Mean-Square-Error (RMSE) statistics for the twelve photo-z codes for both PIT and $\hat{N}(z)$ distributions.}\label{tab:rmse}
\begin{tabular}{lcc}
\hline
\hline
Root-Mean-Square-Error \\
(RMSE) statistics \\
\hline
Photo-z Code & PIT RMSE & N(z) RMSE\\
\hline
\textsc{ANNz2}      & 0.019 & 0.0054\\
\textsc{BPZ}        & 0.032 & 0.0050\\
\textsc{Delight}    & 0.111 & 0.0056\\
\textsc{EAZY}       & 0.054 & 0.0102\\
\textsc{FlexZBoost} & 0.021 & 0.0022\\
\textsc{GPz}        & 0.027 & 0.0042\\
\textsc{LePhare}    & 0.028 & 0.0062\\
\textsc{METAPhoR}   & 0.064 & 0.0081\\
\textsc{CMNN}         & 0.108 & 0.0009\\
\textsc{Skynet}     & 0.054 & 0.0144\\
\textsc{TPZ}        & 0.082 & 0.0031\\
\hline
\trainz		& 0.0025 & 0.0013\\
\end{tabular}
\end{center}
\end{table}
